VOI-aware MCTS

### 2012-02-26 paper/vago.ecai/paper.tex
making it shorter

\begin{abstract}
Upper bounds for the VOI are provided for pure exploration in the
Multi-armed Bandit Problem. A sampling policy based on the upper
bounds is suggested. Empirical evaluation of the policy and
comparison to the UCB1 and UCT policies is provided
on random problem instances as well as on the Go game.
\end{abstract}

@@

\subsection{VOI-based Sample Allocation}

@@

The experiments compare the UCT algorithm \cite{Kocsis.uct} with
a modified version of UCT in which the samples at the first step are
selected according to their VOI estimates.

@@

The sampling policies are first compared on random Multi-armed bandit 
problem instance. Figure~\ref{fig:random-instances} shows experiment results for
randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range $[0,
  1]$, for a range of sample budgets $32..2048$, with multiplicative
step of $2$. The experiment for each number of samples was repeated
20000 times. UCT is always considerably worse than the VOI-aware sampling policy.

@@

One search domain in which Monte-Carlo tree search based on the UCT
sampling policy has been particularly successful is playing the game Go
\cite{Gelly.mogo}. This series of experiments compares
the winning rate of UCT against VOI-aware policies (VCT, ECT, BCT)
in Go and investigates the effect of early stopping and sample
redistribution.

A modified version of Pachi \cite{Braudis.pachi}, a state of the art
Go program, was used for the experiments. The UCT engine was extended
with a VOI-aware sampling policy, and a time allocation mode ensuring
that both the original UCT policy and the VOI-aware policy use the
same average number of samples per node was added. While the UCT
engine is not the most powerful engine of Pachi, it is still a strong
player. On the other hand, additional features of more advanced
engines would obstruct the Monte-Carlo sampling phenomena which are
the subject of the experiment.
\begin{figure}[t]
\centering
\includegraphics[scale=0.5]{vct-wins.pdf}
\caption{Go: winning rate --- UCT against VCT}
\label{fig:vct-against-uct}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
15000 samples (game simulations) per ply, each experiment was repeated
1000 times. Figure~\ref{fig:vct-against-uct}
shows the winning rate of VOI+UCT against UCT vs. the number of
samples. For most numbers of samples per node, VO+UCT outperforms UCT.

>>>>

\begin{abstract}
Upper bounds for the VOI are provided for pure exploration in the
MCTS. A sampling policy based on the upper bounds is
suggested. Empirical evaluation of the policy and comparison to UCB1
and UCT is performed on random bandit instances as well as on Computer
Go.
\end{abstract}

@@

\section{VOI-based Sample Allocation}

@@

@@

The sampling policies are first compared on random Multi-armed bandit 
problem instances. Figure~\ref{fig:random-instances} shows results for
randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range $[0,
  1]$, for a range of sample budgets $32..1024$, with multiplicative
step of $2$. The experiment for each number of samples was repeated
10000 times. UCB1 is always considerably worse than the
VOI-aware sampling policy.

@@

The policies were also compared on Computer Go, a  search domain
in which UCT-based MCTS has been particularly successful
\cite{Gelly.mogo}. A modified version of Pachi \cite{Braudis.pachi}, a state of the art
Go program, was used for the experiments. The UCT engine was extended
with a VOI-aware sampling policy, and a time allocation mode ensuring
that both the original UCT policy and the VOI-aware policy use the
same average number of samples per node was added. (While the UCT
engine is not the most powerful engine of Pachi, it is still a strong
player; on the other hand, additional features of more advanced
engines would obstruct the MCTS phenomena which are the subject of
the experiment.)
\begin{figure}[t]
\centering
\includegraphics[scale=0.5]{vct-wins.pdf}
\caption{Go: winning rate --- UCT against VCT}
\label{fig:vct-against-uct}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
15000 samples per ply, each experiment was repeated
1000 times. Figure~\ref{fig:vct-against-uct}
shows the winning rate of VOI+UCT against UCT vs. the number of
samples. For most numbers of samples per node, VOI+UCT outperforms UCT.
